% la-lab-02-lineqs.tex

\documentclass[11pt]{article}
\usepackage{enumerate}
\usepackage{syllogism} 
\usepackage{october}
\usepackage[table]{xcolor}

\newcounter{aufg}
\setcounter{aufg}{0}
\newcommand{\aufgabe}[0]{\refstepcounter{aufg}\textbf{(\arabic{aufg})}}

\begin{document}

\textbf{Linear Equations}

{\aufgabe} A triangle has vertices $(1,2),(3,–4),(–2,3)$. Find the
area of the triangle.

% Then the area of the triangle is 8 square units.

{\aufgabe} The currents running through an electrical system are given
by the following system of equations. The three currents
$I_{1},I_{2},I_{3}$ are measured in amps. Solve the system to find the
currents in this circuit. (Use elementary row operations to find the
inverse matrix.)
  \begin{align}
    \label{eq:oemeyeib}
    \begin{array}{clclclc}
    I_{1} & + & 2I_{2} & - & I_{3} & = & 0.425 \\
    3I_{1} & - & I_{2} & + & 2I_{3} & = & 2.225 \\
    5I_{1} & + & I_{2} & + & 2I_{3} & = & 3.775
    \end{array}
  \end{align}

  {\aufgabe} Find the equation of the parabola $y=ax^{2}+bx+c$ that
  passes through the following three points: $(-2,40),(1,7),(3,15)$.
  (Use the determinant/adjugate method to find the inverse matrix.)

  % {\aufgabe} Billy's Restaurant ordered 200 flowers for Mother's Day.
  % They ordered carnations at \$1.50 each, roses at \$5.75 each, and
  % daisies at \$2.60 each. They ordered mostly carnations; and 20 fewer
  % roses than daisies. The total order came to \$589.50. How many of
  % each type of flower was ordered?

  {\aufgabe} The Arcadium arcade in Lynchburg, Tennessee uses 3
  different colored tokens for their game machines. For \$20 you can
  purchase any of the following mixtures of tokens: 14 gold, 20
  silver, and 24 bronze; OR, 20 gold, 15 silver, and 19 bronze; OR, 30
  gold, 5 silver, and 13 bronze. What is the monetary value of each
  token?

  {\aufgabe} In the position function for vertical height
  \begin{equation}
    \label{eq:eicheuth}
    s(t)=\frac{1}{2}at^{2}+v_{0}t+s_{0}
  \end{equation}
  $s(t)$ represents height in meters and $t$ represents time in
  seconds.
  \begin{enumerate}
  \item Find the position function for a volleyball served at an
    initial height of one meter, with height of 6.275 meters 0.5
    seconds after serve, and height of 9.1 meters one second after
    serve.
  \item How long until the ball hits the ground on the other side of
    the net if everyone on that team completely misses it?
  \end{enumerate}

  {\aufgabe} Last Tuesday, Regal Cinemas sold a total of 8500 movie
  tickets. Proceeds totaled \$64,600. Tickets can be bought in one of
  3 ways: a matinee admission costs \$5, student admission is \$6 all
  day, and regular admissions are \$8.50. How many of each type of
  ticket was sold if twice as many student tickets were sold as
  matinee tickets?

\end{document}

